\chapter{Abstract}

\begin{english} %switch to English language rules
The following paper provides an insight into various alternative input and output methodes, compares all the individual systems with each other and describes their application areas.
\newline \newline
There are various possibilities how a user can interact with a system without using their hands. There is the option to interact using voice control, eye control, gesture control, muscle control or control based on brain activity. In case of speech control the spoken words are first saved digitally. Later, they are compared with existing patterns and word orders to understand the content and required action of the user. When a user interacts with an eye control system, the eye movement will be tracked by one or more cameras and a software translates these movements into real time mouse movements. Antoher method is the gesture control, which can be divided into several subcategories. Summarizing, they all use a joystick-like element for the direction of movement and additional keys help with the selection of desired symbols or elements. Electromyographic data (EMG) is providing the basis for muscle control, which enable the user to achieve the desired interaction through targeted muscle contraction and relaxation. The control by brain activity utilizes the measurement of the human brain waves with the help of an electroencephalograph (EEG). Increased brain activity in the specific areas can be converted into control commands or into an interaction in general.
\newline \newline
A computer can give feedback either through audio, hapitc or visual output. However, these output methods are not nessecarily tied to specific input methods. Therefore, there are not only systems controled by speech, which provide acoustic feedback, but also muscle control sytems for instance.
\newline \newline
As a conclusion alternative interaction options are used in everyday situations, at work and as assistive technologies. Due to the current popularity of Amazon Alexa and other language assistants, the language output was defined as the most popular type of alternative interaction systems at the moment. On the other hand, the control of brain activity promises the most potential.
\end{english}


